---
title: "School Shooting Projections"
author: "Bianca Doone"
date: "April 3, 2018"
output: pdf_document
---

###Questions:
Are the frequency of shootings really increasing?

Whare are the projections for shootings in the next 10 years? How many shootings can we expect this year? 

This analysis looks at data from 1991 - 2017. We will construct projections for 2018 - 2030 based on the last 20, 10, and 5 years respectively, and compare how the projections change based on different amounts of previous data.


\pagebreak

```{r, echo=FALSE, warning = FALSE, message=FALSE}
library(ggplot2)
library(lubridate)
library(R2jags)
library(rjags)
```

```{r,  echo=FALSE}
ss_final = read.csv("C:/Users/bdoone/Desktop/ss_final.csv", header = T)
shootings2 = read.csv("C:/Users/bdoone/Desktop/shootings2.csv", header = T)
year_df = read.csv("C:/Users/bdoone/Desktop/year_df.csv", header = T)
```


```{r, echo=FALSE}
#create difference variable
shootings2$daysfromlast <- rep(NA, length(shootings2$Days))
shootings2$daysfromlast[1] = 0

for(i in 2:length(shootings2$Days)){
  shootings2$daysfromlast[i] = shootings2$Day[i] - shootings2$Days[i-1]
}


```


#Summary Statistics (TOTAL)

```{r}
mean(year_df$Shootings)
table(ss_final$Shooting)
year_df$Prop.Shootings.Year

#add table year and number shootings/proportion
```


#Incidence of Shooting Over Time
```{r}
#basic histogram - number of shootings per year (with proportions table)
ggplot(data=shootings2, aes(x = as.factor(Year), fill = as.factor(Year))) +
  geom_bar()

print(year_df$Prop.Shootings.Year)

#basic graph - days from last shooting against shooting (with table maximums)
ggplot(data = shootings2, aes(x = Days, y = daysfromlast)) + geom_line()

print(year_df$Max.difference)
```

#My Model

Shooting model (by year):



$$y_t | p_t, r \sim \mbox{NB}(p_t,r) $$
where $p_t = \frac{r}{r + \lambda_t}$ and $\lambda_t = \mbox{e}^{\mu_t}$

$$\mu_t = \beta_0 + \beta_1 t + \mbox{AR}(1) $$
where $\mbox{AR}(1) = u_t = \rho u_{t-1} + \varepsilon_t$

With vague priors:

$$\beta_0 \sim N(0, 100^2), \beta_1 \sim N(0, 100^2), r \sim U(0,50)$$
$$\rho \sim U(-1,1), \varepsilon \sim N(0,10) $$

$y_t$ is the number of shootings in year $t$.

```{r, echo=FALSE}
#timeseries, all data

jags.data.t <- list(y.t = year_df$Shootings[c(1:27)], 
                  x.t = year_df$Year[c(1:27)], 
                  nyears=length(year_df$Year[c(1:27)]))
parnames.t <- c("lambda.t", "p.t", "yhat.t")
set.seed(1)

model.t <- "model{
y.t[1] ~ dnegbin(p.t[1],r)
yhat.t[1] ~ dnegbin(p.t[1],r)
p.t[1] <- r/(r+lambda.t[1])
lambda.t[1] <- exp(mu.t[1])
mu.t[1] <- beta0 + beta1*x.t[1] + u.t[1]
u.t[1] ~ dnorm(0, tau.u)
tau.u <- (1-pow(rho,2))/pow(sigma,2)

for(t in 2:nyears){
y.t[t] ~ dnegbin(p.t[t],r)
yhat.t[t] ~ dnegbin(p.t[t],r)
p.t[t] <- r/(r+lambda.t[t])
lambda.t[t] <- exp(mu.t[t])
mu.t[t] <- beta0 + beta1*x.t[t] + u.t[t]
u.t[t] <- rho*u.t[t-1]
}
beta0 ~ dnorm(0,1/100^2)
beta1 ~ dnorm(0, 1/100^2)
r ~ dunif(0,50)
rho ~ dunif(-1,1)
sigma ~ dunif(0,10)
}"

mod.t <-jags(data = jags.data.t, 
                    parameters.to.save= c("y.t", parnames.t),
           n.chains = 3, n.burnin = 1000, n.iter = 1000+1000, n.thin = 1,
           model.file = textConnection(model.t))

y.t.pred.t = mod.t$BUGSoutput$sims.list[["yhat.t"]]

y.t.med.t = rep(0,27)
y.t.low.t = rep(0,27)
y.t.high.t = rep(0,27)

for(i in 1:27){
  y.t.med.t[i] = median(y.t.pred.t[,i])
  y.t.low.t[i] = quantile(y.t.pred.t[,i], c(.025))
  y.t.high.t[i] = quantile(y.t.pred.t[,i], c(.975))
}

plot(year_df$Shootings[c(1:27)]~year_df$Year[c(1:27)], 
     main = "Number of Shootings per Year", 
     ylab = "Shootings", xlab = "Year",
     ylim = c(0,70))
lines(y.t.med.t~year_df$Year[c(1:27)], col = "blue")
lines(y.t.low.t~year_df$Year[c(1:27)], col = "red", lty = 2)
lines(y.t.high.t~year_df$Year[c(1:27)], col = "red", lty = 2)
```




Based on the last 20 years:

```{r, echo=FALSE}
jags.data20 <- list(y.t = year_df$Shootings[c(7:27)], 
                  x.t = year_df$Year[c(7:27)], 
                  nyears=length(year_df$Year[c(7:27)]),
                  x.p = c(2018:2030))
parnames20 <- c("lambda.t", "p.t", "yhat.t")
set.seed(1)

model20 <- "model{
y.t[1] ~ dnegbin(p.t[1],r)
yhat.t[1] ~ dnegbin(p.t[1],r)
p.t[1] <- r/(r+lambda.t[1])
lambda.t[1] <- exp(mu.t[1])
mu.t[1] <- beta0 + beta1*x.t[1] + u.t[1]
u.t[1] ~ dnorm(0, tau.u)
tau.u <- (1-pow(rho,2))/pow(sigma,2)

for(t in 2:nyears){
y.t[t] ~ dnegbin(p.t[t],r)
yhat.t[t] ~ dnegbin(p.t[t],r)
p.t[t] <- r/(r+lambda.t[t])
lambda.t[t] <- exp(mu.t[t])
mu.t[t] <- beta0 + beta1*x.t[t] + u.t[t]
u.t[t] <- rho*u.t[t-1]
}
beta0 ~ dnorm(0,1/100^2)
beta1 ~ dnorm(0, 1/100^2)
r ~ dunif(0,50)
rho ~ dunif(-1,1)
sigma ~ dunif(0,10)

y.p[1] ~ dnegbin(p.p[1],r)
p.p[1] <- r/(r+lambda.p[1])
lambda.p[1] <- exp(mu.p[1])
mu.p[1] <- beta0 + beta1*x.p[1] + u.p[1]
u.p[1] <- rho*u.t[nyears]

for(p in 2:P){
y.p[p] ~ dnegbin(p.p[p],r)
p.p[p] <- r/(r+lambda.p[p])
lambda.p[p] <- exp(mu.p[p])
mu.p[p] <- beta0 + beta1*x.p[p] + u.p[p]
u.p[p] <- rho*u.p[p-1]
}
}"
P = 13
mod20 <-jags(data = c(jags.data20, P=P),
                    parameters.to.save= c("y.t", parnames20, "y.p"),
           n.chains = 3, n.burnin = 1000, n.iter = 1000+1000, n.thin = 1,
           model.file = textConnection(model20))

y.t.20 = mod20$BUGSoutput$sims.list[["yhat.t"]]

y.t.med20 = rep(0,21)
y.t.low20 = rep(0,21)
y.t.high20 = rep(0,21)

for(i in 1:21){
  y.t.med20[i] = median(y.t.20[,i])
  y.t.low20[i] = quantile(y.t.20[,i], c(.025))
  y.t.high20[i] = quantile(y.t.20[,i], c(.975))
}

y.p.20 = mod20$BUGSoutput$sims.list[["y.p"]]

y.p.med20 = rep(0,13)
y.p.low20 = rep(0,13)
y.p.high20 = rep(0,13)

for(i in 1:13){
  y.p.med20[i] = median(y.p.20[,i])
  y.p.low20[i] = quantile(y.p.20[,i], c(.025))
  y.p.high20[i] = quantile(y.p.20[,i], c(.975))
}
spacer = rep("",13)
plot(c(year_df$Shootings[c(1:27)],spacer)~c(1991:2030), 
     main = "Number of Shootings Last 20 Years with Forecast to 2030", 
     ylab = "Shootings", xlab = "Year",
     ylim = c(0,100))
lines(y.t.med20~c(1997:2017), col = "blue")
lines(y.t.low20~c(1997:2017), col = "red", lty = 2)
lines(y.t.high20~c(1997:2017), col = "red", lty = 2)
lines(y.p.med20~c(2018:2030), col = "purple")
lines(y.p.high20~c(2018:2030), col = "purple", lty = 2)
lines(y.p.low20~c(2018:2030), col = "purple", lty = 2)
```


Based on the last 10 years:


```{r, echo=FALSE}
jags.data10 <- list(y.t = year_df$Shootings[c(17:27)], 
                  x.t = year_df$Year[c(17:27)], 
                  nyears=length(year_df$Year[c(17:27)]),
                  x.p = c(2018:2030))
parnames10 <- c("lambda.t", "p.t", "yhat.t")
set.seed(1)

model10 <- "model{
y.t[1] ~ dnegbin(p.t[1],r)
yhat.t[1] ~ dnegbin(p.t[1],r)
p.t[1] <- r/(r+lambda.t[1])
lambda.t[1] <- exp(mu.t[1])
mu.t[1] <- beta0 + beta1*x.t[1] + u.t[1]
u.t[1] ~ dnorm(0, tau.u)
tau.u <- (1-pow(rho,2))/pow(sigma,2)

for(t in 2:nyears){
y.t[t] ~ dnegbin(p.t[t],r)
yhat.t[t] ~ dnegbin(p.t[t],r)
p.t[t] <- r/(r+lambda.t[t])
lambda.t[t] <- exp(mu.t[t])
mu.t[t] <- beta0 + beta1*x.t[t] + u.t[t]
u.t[t] <- rho*u.t[t-1]
}
beta0 ~ dnorm(0,1/100^2)
beta1 ~ dnorm(0, 1/100^2)
r ~ dunif(0,50)
rho ~ dunif(-1,1)
sigma ~ dunif(0,10)

y.p[1] ~ dnegbin(p.p[1],r)
p.p[1] <- r/(r+lambda.p[1])
lambda.p[1] <- exp(mu.p[1])
mu.p[1] <- beta0 + beta1*x.p[1] + u.p[1]
u.p[1] <- rho*u.t[nyears]

for(p in 2:P){
y.p[p] ~ dnegbin(p.p[p],r)
p.p[p] <- r/(r+lambda.p[p])
lambda.p[p] <- exp(mu.p[p])
mu.p[p] <- beta0 + beta1*x.p[p] + u.p[p]
u.p[p] <- rho*u.p[p-1]
}
}"
P = 13
mod10 <-jags(data = c(jags.data10, P=P),
                    parameters.to.save= c("y.t", parnames10, "y.p"),
           n.chains = 3, n.burnin = 1000, n.iter = 1000+1000, n.thin = 1,
           model.file = textConnection(model10))

y.t.10 = mod10$BUGSoutput$sims.list[["yhat.t"]]

y.t.med10 = rep(0,11)
y.t.low10 = rep(0,11)
y.t.high10 = rep(0,11)

for(i in 1:11){
  y.t.med10[i] = median(y.t.10[,i])
  y.t.low10[i] = quantile(y.t.10[,i], c(.025))
  y.t.high10[i] = quantile(y.t.10[,i], c(.975))
}

y.p.10 = mod10$BUGSoutput$sims.list[["y.p"]]

y.p.med10 = rep(0,13)
y.p.low10 = rep(0,13)
y.p.high10 = rep(0,13)

for(i in 1:13){
  y.p.med10[i] = median(y.p.10[,i])
  y.p.low10[i] = quantile(y.p.10[,i], c(.025))
  y.p.high10[i] = quantile(y.p.10[,i], c(.975))
}
spacer = rep("",13)
plot(c(year_df$Shootings[c(1:27)],spacer)~c(1991:2030), 
     main = "Number of Shootings Last 10 Years with Forecast to 2030", 
     ylab = "Shootings", xlab = "Year",
     ylim = c(0,100))
lines(y.t.med10~c(2007:2017), col = "blue")
lines(y.t.low10~c(2007:2017), col = "red", lty = 2)
lines(y.t.high10~c(2007:2017), col = "red", lty = 2)
lines(y.p.med10~c(2018:2030), col = "purple")
lines(y.p.high10~c(2018:2030), col = "purple", lty = 2)
lines(y.p.low10~c(2018:2030), col = "purple", lty = 2)
```


Based on the last 5 years:

```{r, echo=FALSE}
jags.data5 <- list(y.t = year_df$Shootings[c(22:27)], 
                  x.t = year_df$Year[c(22:27)], 
                  nyears=length(year_df$Year[c(22:27)]),
                  x.p = c(2018:2030))
parnames5 <- c("lambda.t", "p.t", "yhat.t")
set.seed(1)

model5 <- "model{
y.t[1] ~ dnegbin(p.t[1],r)
yhat.t[1] ~ dnegbin(p.t[1],r)
p.t[1] <- r/(r+lambda.t[1])
lambda.t[1] <- exp(mu.t[1])
mu.t[1] <- beta0 + beta1*x.t[1] + u.t[1]
u.t[1] ~ dnorm(0, tau.u)
tau.u <- (1-pow(rho,2))/pow(sigma,2)

for(t in 2:nyears){
y.t[t] ~ dnegbin(p.t[t],r)
yhat.t[t] ~ dnegbin(p.t[t],r)
p.t[t] <- r/(r+lambda.t[t])
lambda.t[t] <- exp(mu.t[t])
mu.t[t] <- beta0 + beta1*x.t[t] + u.t[t]
u.t[t] <- rho*u.t[t-1]
}
beta0 ~ dnorm(0,1/100^2)
beta1 ~ dnorm(0, 1/100^2)
r ~ dunif(0,50)
rho ~ dunif(-1,1)
sigma ~ dunif(0,10)

y.p[1] ~ dnegbin(p.p[1],r)
p.p[1] <- r/(r+lambda.p[1])
lambda.p[1] <- exp(mu.p[1])
mu.p[1] <- beta0 + beta1*x.p[1] + u.p[1]
u.p[1] <- rho*u.t[nyears]

for(p in 2:P){
y.p[p] ~ dnegbin(p.p[p],r)
p.p[p] <- r/(r+lambda.p[p])
lambda.p[p] <- exp(mu.p[p])
mu.p[p] <- beta0 + beta1*x.p[p] + u.p[p]
u.p[p] <- rho*u.p[p-1]
}
}"
P = 13
mod5 <-jags(data = c(jags.data5, P=P),
                    parameters.to.save= c("y.t", parnames5, "y.p"),
           n.chains = 3, n.burnin = 1000, n.iter = 1000+1000, n.thin = 1,
           model.file = textConnection(model5))

y.t.5 = mod5$BUGSoutput$sims.list[["yhat.t"]]

y.t.med5 = rep(0,6)
y.t.low5 = rep(0,6)
y.t.high5 = rep(0,6)

for(i in 1:6){
  y.t.med5[i] = median(y.t.5[,i])
  y.t.low5[i] = quantile(y.t.5[,i], c(.025))
  y.t.high5[i] = quantile(y.t.5[,i], c(.975))
}

y.p.5 = mod5$BUGSoutput$sims.list[["y.p"]]

y.p.med5 = rep(0,13)
y.p.low5 = rep(0,13)
y.p.high5 = rep(0,13)

for(i in 1:13){
  y.p.med5[i] = median(y.p.5[,i])
  y.p.low5[i] = quantile(y.p.5[,i], c(.025))
  y.p.high5[i] = quantile(y.p.5[,i], c(.975))
}
spacer = rep("",13)
plot(c(year_df$Shootings[c(1:27)],spacer)~c(1991:2030), 
     main = "Number of Shootings Last 5 Years with Forecast to 2030", 
     ylab = "Shootings", xlab = "Year",
     ylim = c(0,100))
lines(y.t.med5~c(2012:2017), col = "blue")
lines(y.t.low5~c(2012:2017), col = "red", lty = 2)
lines(y.t.high5~c(2012:2017), col = "red", lty = 2)
lines(y.p.med5~c(2018:2030), col = "purple")
lines(y.p.high5~c(2018:2030), col = "purple", lty = 2)
lines(y.p.low5~c(2018:2030), col = "purple", lty = 2)
```


2018: Projected Median number of shootings, 20, 10, 5 years data w/ 95% CI

```{r, echo=FALSE}
pe18 = data.frame(matrix(c(quantile(y.p.20[1,], c(.025,.5,.975)),
                           quantile(y.p.10[1,], c(.025,.5,.975)),
                           quantile(y.p.5[1,], c(.025,.5,.975))),
                         nrow = 3, byrow = T))
pe20 = data.frame(matrix(c(quantile(y.p.20[3,], c(.025,.5,.975)),
                           quantile(y.p.10[3,], c(.025,.5,.975)),
                           quantile(y.p.5[3,], c(.025,.5,.975))),
                         nrow = 3, byrow = T))
pe25 = data.frame(matrix(c(quantile(y.p.20[8,], c(.025,.5,.975)),
                           quantile(y.p.10[8,], c(.025,.5,.975)),
                           quantile(y.p.5[8,], c(.025,.5,.975))),
                         nrow = 3, byrow = T))
pe30 = data.frame(matrix(c(quantile(y.p.20[13,], c(.025,.5,.975)),
                           quantile(y.p.10[13,], c(.025,.5,.975)),
                           quantile(y.p.5[13,], c(.025,.5,.975))),
                         nrow = 3, byrow = T))

colnames(pe18) = c("2.5%", "50%", "97.5%")
rownames(pe18) = c("20 Years", "10 Years", "5 Years")
colnames(pe20) = c("2.5%", "50%", "97.5%")
rownames(pe20) = c("20 Years", "10 Years", "5 Years")
colnames(pe25) = c("2.5%", "50%", "97.5%")
rownames(pe25) = c("20 Years", "10 Years", "5 Years")
colnames(pe30) = c("2.5%", "50%", "97.5%")
rownames(pe30) = c("20 Years", "10 Years", "5 Years")

pe18
```

2020: Projected Median number of shootings, 20, 10, 5 years data w/ 95% CI
```{r, echo=FALSE}
pe20
```

2025: Projected Median number of shootings, 20, 10, 5 years data w/ 95% CI
```{r, echo=FALSE}
pe25
```

2030: Projected Median number of shootings, 20, 10, 5 years data w/ 95% CI
```{r, echo=FALSE}
pe30
```

















