---
title: "Bayes Project"
author: "Bianca, Heather, Casey"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(readr)
```


```{r combining csvs}
shootingsafter <- read.csv("~/ShootingsAfter1990.csv")
moreshootings <- read.csv("~/Shootings2016.csv")

write.csv(rbind(shootingsafter,moreshootings), "AllShootings.csv")

allshootings <- read.csv("~/AllShootings.csv")
```

### Intro

School shootings have long been unique in America for their frequency and intensity. We have obtained a dataset on all US school shootings since 1766 through UMass GRID in order to begin research on an issue that has had very little scientific study. By examining this data-set in a statistically rigorous way, we hope to gain a better understanding of a tragic phenomenon that does not seem to be in decline. 


```{r plots, echo=FALSE}
dates <- allshootings$Date
dates <- dates[2:length(dates)]

allshootings$year <- format(as.Date(allshootings$Date,format="%m/%d/%Y"),"%Y")

allshootings$victims <- allshootings$Victim.s..Deceased..at.school.+allshootings$Victim.s..Deceased..off.campus.

library(ggplot2)

df_ <- data.frame(id = 1:length(allshootings$Victim.s..Deceased..at.school.))

ggplot(df_,aes(x=allshootings$Weapon.s..Categories,y = allshootings$Victim.s..Deceased..at.school.)) + geom_point()   + labs(title = "Number of Victims by Gun Type")

ggplot(df_,aes(x=allshootings$year,y = allshootings$Victim.s..Deceased..at.school.)) + geom_point() + labs(title = "Number of Victims over Time")  + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```




Potential model for number of victims in an incident:
Let
$$y_i = \text{ number of victims in incident }$$
$$r_i = \text{ race of the perpetrator  }$$
$$g_i = \text{ gun used by perpetrator} $$
$$s_i = \text{ type of institution} $$

$$\alpha_j[i] = \text{ random intercept accounting for geographic region}$$

$$Y_i \sim Pois(\lambda_i)$$
$$log(\lambda_i) = \alpha_{j[i]} + \beta_1*r_i + \beta_2*g_i +  \beta_3*t_i + \beta_4*s_i$$


```{r}
library(rjags)
library(R2jags)

model <- "
model {
for( i in 1:n) {	
log(lambda.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0
  + b1*r[i] + b2*gt[i]
  )
  y.i[i] ~ dpois(lambda.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)

} # end model
"
jags.data <- list(y.i = df_$victims, 
                   r = as.factor(df_$r), gt =as.factor(df_$gt),  n=length(df_$victims))

parnames <- c( "b1", "b0","b2")
mod <- jags(data = jags.data, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))
# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod$BUGSoutput$sims.array
par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,3))
hist(c(mcmc.array[,,"b0"]), freq = F, main = "", xlab ="Intercept")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)

hist(c(mcmc.array[,,"b1"]), freq = F, main = "", xlab ="Race")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)

hist(c(mcmc.array[,,"b1"]), freq = F, main = "", xlab ="Gun Type ")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)

```



 
