---
title: "Bayes Project"
author: "Bianca, Heather, Casey"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(googledrive)
library(rjags)
library(R2jags)
```


```{r reading in data combining csvs, warning=FALSE, results='hide'}
gun_regulations <- read.csv("~/gun_regulation.csv")
shootingsafter <- read.csv("~/ShootingsAfter1991.csv")
moreshootings <- read.csv("~/Shootings2016.csv")

write.csv(rbind(shootingsafter,moreshootings), "AllShootings.csv")

allshootings <- read.csv("~/AllShootings.csv")
```

### Intro

School shootings have long been unique in America for their frequency and intensity. We have obtained a dataset on all US school shootings since 1766 through UMass GRID in order to begin research on an issue that has had very little scientific study. By examining this data-set in a statistically rigorous way, we hope to gain a better understanding of a tragic phenomenon that does not seem to be in decline. 


```{r plots, echo=FALSE}
dates <- allshootings$Date
dates <- dates[2:length(dates)]

allshootings$Date <- as.Date(allshootings$Date,format="%m/%d/%Y")
allshootings$year <- format(allshootings$Date,"%Y")

number <- rep(NA, length(allshootings$Victim.s..Deceased..at.school.))

for (i in 1:length(allshootings$Victim.s..Deceased..at.school.)){
  if (is.na(allshootings$Victim.s..Deceased..at.school.)[i]==TRUE){
    number[i] <- 0
  }
    else{
      number[i]<- allshootings$Victim.s..Deceased..at.school.[i]
  }
}

#allshootings$victims <- allshootings$Victim.s..Deceased..at.school.+allshootings$Victim.s..Deceased..off.campus.

ggplot(gun_regulations,aes(x=gun_regulations$state,y = gun_regulations$lawtotal)) + geom_point()   + labs(title = "Total Number of Gun Regulations by State")

ggplot(allshootings,aes(x=allshootings$Weapon.s..Categories,y = allshootings$Victim.s..Deceased..at.school.)) + geom_point()   + labs(title = "Number of Victims by Gun Type", y="Number of Victims", x="Type of Gun")

ggplot(allshootings, aes(x=allshootings$year,y = allshootings$Victim.s..Deceased..at.school.)) + geom_point() + labs(title = "Number of Victims over Time", y="Number of Victims", x="Year")  + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Potential model for number of shootings:
Let
$$y_{j} = \text{ number of shootings in state } j $$
$$ m_j = \text{ mental health resourses or mental health prevalence in state } j$$

$$g.reg_{j} = \text{ gun regulations} $$

$$\alpha_{k[j]} = \text{ random intercept accounting for geographic region}$$

$$Y_{j} \sim Pois(\lambda_{j})$$
$$log(\lambda_{j}) = \alpha_{k} + \beta_1*g.reg_{j} + + \beta_2*m_{j} + \beta_3*time  $$


Potential model for gun type used: 

$$ \text{ gun type } = G_{i} \sim Multinomial(J, p_1, ..., p_J) $$
$$log(\frac {P(G_i = j|g.reg_i)}{P(G_i = J|g.reg_i)}) = \alpha_j + \beta_j*g.reg_i  $$
$$ j = 1, ..., J-1 $$
$$ J \text{ is the reference gun type} $$



Potential model for number of victims in an incident:
Let
$$y_i = \text{ number of victims in incident }$$
$$r_i = \text{ race of the perpetrator  }$$
$$g_i = \text{ gun used by perpetrator} $$
$$s_i = \text{ type of institution} $$

$$\alpha_j[i] = \text{ random intercept accounting for geographic region}$$

$$Y_i \sim Pois(\lambda_i)$$
$$log(\lambda_i) = \alpha_{j[i]} + \beta_1*r_i + \beta_2*g_i +  \beta_3*t_i + \beta_4*s_i$$


```{r, results='hide'}
model <- "
model {
for( i in 1:n) {	
log(lambda.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0
  + b1*r[i] + b2*gt[i]
  )
  y.i[i] ~ dpois(lambda.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)

} # end model
"
jags.data <- list(y.i = allshootings$Victim.s..Deceased..at.school., 
                   r = allshootings$Shooter.s..or.Attacker.s..Race, gt =allshootings$Weapon.s..Categories,  n=length(allshootings$Victim.s..Deceased..at.school.))

parnames <- c( "b1", "b0","b2")
mod <- jags(data = jags.data, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))
```

```{r}
# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod$BUGSoutput$sims.array
par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,3))
hist(c(mcmc.array[,,"b0"]), freq = F, main = "", xlab ="Intercept")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)

hist(c(mcmc.array[,,"b1"]), freq = F, main = "", xlab ="Race")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)

hist(c(mcmc.array[,,"b1"]), freq = F, main = "", xlab ="Gun Type ")
curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
```



 
