---
title: "Bayes Project"
author: "Bianca, Heather, Casey"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Intro

School shootings have long been unique in America for their frequency and intensity. We have obtained a dataset on all US school shootings since 1766 through UMass GRID in order to begin research on an issue that has had very little scientific study. By examining this data-set in a statistically rigorous way, we hope to gain a better understanding of a tragic phenomenon that does not seem to be in decline. 


### Data
```{r}
csv.data <- read.csv("data.csv")
head(colnames(csv.data),10)
```


Potential model:
Let
$$y_i = \text{ length of sentence of the ith school shooting perpetrator}$$
$$r_i = \text{ race of the perpetrator  }$$
$$g_i = \text{ gun used by perpetrator} $$
$$s_i = \text{ type of institution} $$

$$\alpha_j[i] = \text{ random intercept accounting for geographic region}$$

$$Y_i \sim Pois(\lambda_i)$$
$$log(\lambda_i) = \alpha_{j[i]} + \beta_1*r_i + \beta_2*g_i +  \beta_3*t_i + \beta_4*s_i$$

We will also look at interactions and possible transformations of covariates as needed. 

Questions:  
 - Would a time series component work for this model?  
 - Should institution type be a hierarchical structure?  
 
