---
title: "Bayes Project (Casey's Part)"
author: "Bianca, Heather, Casey"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(rjags)
library(R2jags)
```

## Data

```{r reading in data combining csvs,echo=FALSE, warning=FALSE, results='hide'}

#setwd("/Users/gcgibson/BayesProject/")

shootingsafter <- read.csv("/Users/gcgibson/BayesProject/ShootingsAfter1991.csv")
moreshootings <- read.csv("/Users/gcgibson/BayesProject/Shootings2016.csv")

write.csv(rbind(shootingsafter,moreshootings), "AllShootings.csv")

allshootings <- read.csv("/Users/gcgibson/BayesProject/AllShootings.csv")

ames.data <- allshootings$Victim.s..Deceased..at.school.

number_of_victims <-c()
s <- c()
for (i in 1:length(ames.data)){
  if (is.na(ames.data[i])){
    #print ("hello")
    number_of_victims <- c(number_of_victims,0)
  } else if (ames.data[i] == "None"){
    number_of_victims <- c(number_of_victims,0)
  } else{
    number_of_victims <- c(number_of_victims,ames.data[i])
  }
}

allshootings$number_of_victims <- number_of_victims
plot(1:length(number_of_victims),number_of_victims)
#ggplot(allshootings, aes(x=allshootings$Date,y = allshootings$number_of_victims))  geom_point() + labs(title = "Number of Victims over Time", y="Number of Victims", x="Year")  + theme(axis.text.x = element_text(angle = 90, hjust = 1))


type_of_gun <- matrix(0,nrow=length(number_of_victims),ncol=4)
tmp <- allshootings$Weapon.s..Categories

for (i in 1:length(tmp)){
  if (tmp[i] == "Handgun"){
    type_of_gun[i,1] =  1
  } else if (tmp[i] == "Rifle"){
    type_of_gun[i,2] =  1
  }  else if (tmp[i] == "Shotgun"){
    type_of_gun[i,3] =  1
  }  else {
    type_of_gun[i,4] =  1
  }
}
```


We first attempt a simple Poisson regression. 
$$Y_t = \text{number of victims of the t'th school shooting}$$
$$ X_{1:4} = \text{ gun type encoded as }$$
$$\text{ 0 = handgun}$$
$$\text{ 1 = rifle}$$
$$\text{ 2 = shotgun}$$
$$\text{ 3 = unknown/other}$$

$$Y_t \sim Poisson(\lambda)$$
$$log(\lambda) = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta3x_3$$

### Poisson


```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
for(i in 1:N){
      y[i] ~ dpois(lambda[i])
      log(lambda[i]) <- mu[i]
      mu[i] <- beta4 + beta1*x1[i] +beta2*x2[i] + beta3*x3[i]    
      }     
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```

We see that neither handgun nor rifle has a significant effect on the number of victims. 

What if we control for race and age?

$$log(\lambda) = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta3x_3 + \beta_5r + \beta_6a$$
### Poisson w Covariates

```{r,echo=FALSE}
race <- allshootings$Shooter.s..or.Attacker.s..Race
race_clean <- c()
for (i in 1:length(race)){
  if (race[i] == "African American"){
    race_clean <- c(race_clean,0)
  } else if (race[i] == "Caucasian"){
    race_clean <- c(race_clean,1)
  }  else if (race[i] == "Hispanic"){
    race_clean <- c(race_clean,2)
  }   else{
    race_clean <- c(race_clean,4)
  }
}
age <- allshootings$Shooter.s..or.Attacker.s..Age
age_clean <- c()
for (i in 1:length(age)){
  if ( 0 < as.numeric(age[i]) & as.numeric(age[i]) <10 ){
    age_clean <- c(age_clean,0)
  } else if (10 < as.numeric(age[i]) & as.numeric(age[i]) < 20){
    age_clean <- c(age_clean,1)
  }   else{
    age_clean <- c(age_clean,2)
  }
}
```




```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
for(i in 1:N){
      y[i] ~ dpois(lambda[i])
      log(lambda[i]) <- mu[i]
      mu[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      }     
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 
}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```


We still don't see any effect of rifle on number of victims, but we see a negative effect of a handgun. That is, we can't say that rifles kill more people than other weapons categories (what I hoped we would find), but we can say that handguns kill fewer people than other weapons. From a policy perspective it makes sense to limit weapons to handguns. 


What happens if we control for the large number of zeros present in the data?

$$Y_t \sim \begin{cases}\pi_t +(1-\pi_t)\cdot e^{\mu_t} & \text{ if } \ y_t=0 \\
(1-\pi_t)\cdot Poisson(\mu_t)  & \text{ if } \ y_t >0 \end{cases}$$
$$logit(\pi_t) =\alpha_0 + \alpha_1x_1 + \alpha_2x_2 + \alpha3x_3 + \alpha_5r + \alpha_6a$$ 
$$log(\lambda_t) =\beta_0 + \beta_1x_1 + \beta_2x_2 + \beta3x_3 + \beta_5r + \beta_6a$$ 
### Zero Inflated Poisson w Covariates

```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
    for(i in 1:N){
      y[i] ~ dpois(lambda.hacked[i])
      lambda.hacked[i] <- lambda[i]*(1-zero[i]) + 1e-10*zero[i]
      lambda[i] <- exp(mu.count[i])
      mu.count[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      
      ## Zero-Inflation
      zero[i] ~ dbern(pi[i])
      pi[i] <- ilogit(mu.binary[i])
      mu.binary[i] <- alpha1*x1[i] +alpha2*x2[i] + alpha3*x3[i] + alpha4  + alpha5*race[i] + alpha6*age[i]
    }  
    for(i in 1:N){
      pp[i] ~ dpois(plambda.hacked[i])
      plambda.hacked[i] <- plambda[i]*(1-zero[i]) + 1e-10*zero[i]
      plambda[i] <- exp(pmu.count[i])
      pmu.count[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      
      ## Zero-Inflation
      pzero[i] ~ dbern(ppi[i])
      ppi[i] <- ilogit(pmu.binary[i])
      pmu.binary[i] <- alpha1*x1[i] +alpha2*x2[i] + alpha3*x3[i] + alpha4  + alpha5*race[i] + alpha6*age[i]
    }  
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 

    alpha1 ~ dnorm(mu.beta,tau.beta) 
    alpha2 ~ dnorm(mu.beta,tau.beta) 
    alpha3 ~ dnorm(mu.beta,tau.beta) 
    alpha4 ~ dnorm(mu.beta,tau.beta) 
    alpha5 ~ dnorm(mu.beta,tau.beta) 
    alpha6 ~ dnorm(mu.beta,tau.beta) 

}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y, 
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6","alpha1","alpha2","alpha3","alpha4","alpha5","alpha6","pp")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```


Ah-ha! if we use the Zero-inflated model we see that rifles do have a positive association with a higher number of victims. This makes sense because the effect of the weapon only matters if the shooter is able to use it (which we use 0 victims as a proxy).



Because we are handling time-series data, errors may be correlated across time. In order to check this, we examine the residuals of the ZIP model fit.

```{r}
ppd <- c()
for (i in 1:length(dat$x.1)){
    ppd <- c(ppd,mcmc.array[,,paste(paste("pp[",i,sep=""),"]",sep="")][1]) 
}

plot(ppd-dat$y)
```

The residuals do look nice and centered around 0, so I'm not sure it is really necessary to to use an AR 
error? We can fit an AR(1) process to the residuals and see if it gives a reasonable fit.



### Poisson + Ar with Covariates

```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
mu.count[1] <-  beta1*x1[1] +beta2*x2[1] + beta3*x3[1] + beta4  + beta5*race[1] + beta6*age[1]
for(i in 2:N){
      y[i] ~ dpois(lambda[i])
      lambda[i] <- exp(mu[i])
      mu[i] <- mu.count[i] + ar1 * ( y[i-1] - mu.count[i-1] )
      mu.count[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      
      
    }  
    ## Priors 
    ar1 ~ dunif(-1.1,1.1) 
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 

 

}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6","alpha1","alpha2","alpha3","alpha4","alpha5","alpha6")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```


### Hierarchical Extension

In order to take the hierarchical data structure into account, we can cluster by region

```{r}

allshootings$region[allshootings$State == "CT" | 
                allshootings$State == "ME"| 
                allshootings$State == "MA"| 
                allshootings$State == "NH"| 
                allshootings$State == "RI"| 
                allshootings$State == "VT"| 
                allshootings$State == "NJ"| 
                allshootings$State == "NY"| 
                allshootings$State == "PA"| 
                allshootings$State == "NY "] <- 1

allshootings$region[allshootings$State == "IL" | 
                allshootings$State == "IN"| 
                allshootings$State == "MI"| 
                allshootings$State == "OH"| 
                allshootings$State == "WI"| 
                allshootings$State == "IA"| 
                allshootings$State == "KS"| 
                allshootings$State == "MN"| 
                allshootings$State == "MO"| 
                allshootings$State == "NE"| 
                allshootings$State == "ND"| 
                allshootings$State == "SD"] <- 2

allshootings$region[allshootings$State == "DE" | 
                allshootings$State == "FL"| 
                allshootings$State == "GA"| 
                allshootings$State == "MD"| 
                allshootings$State == "NC"| 
                allshootings$State == "SC"| 
                allshootings$State == "VA"| 
                allshootings$State == "Wash D.C"| 
                allshootings$State == "WV"| 
                allshootings$State == "AL"| 
                allshootings$State == "KY"| 
                allshootings$State == "MS"| 
                allshootings$State == "TN"| 
                allshootings$State == "AR"| 
                allshootings$State == "LA"| 
                allshootings$State == "OK"| 
                allshootings$State == "TX"| 
                allshootings$State == "FL "] <- 3

allshootings$region[allshootings$State == "AZ" | 
                allshootings$State == "CO"| 
                allshootings$State == "ID"| 
                allshootings$State == "MT"| 
                allshootings$State == "NV"| 
                allshootings$State == "NM"| 
                allshootings$State == "UT"| 
                allshootings$State == "WY"| 
                allshootings$State == "AK"| 
                allshootings$State == "CA"| 
                allshootings$State == "HI"| 
                allshootings$State == "OR"| 
                allshootings$State == "WA"] <- 4

```

We first include a random intercept term.

$$log(\lambda_i) = \alpha_{j[i]} + same above$$

```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
    for(i in 1:N){
      y[i] ~ dpois(lambda.hacked[i])
      lambda.hacked[i] <- lambda[i]*(1-zero[i]) + 1e-10*zero[i]
      lambda[i] <- exp(mu.count[i])
      mu.count[i] <- beta.j[region[i]] + beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta5*race[i] + beta6*age[i]
      
      ## Zero-Inflation
      zero[i] ~ dbern(pi[i])
      pi[i] <- ilogit(mu.binary[i])
      mu.binary[i] <- alpha.j[region[i]] + alpha1*x1[i] +alpha2*x2[i] + alpha3*x3[i] +  alpha5*race[i] + alpha6*age[i]
    }  
    for(i in 1:N){
      pp[i] ~ dpois(plambda.hacked[i])
      plambda.hacked[i] <- plambda[i]*(1-zero[i]) + 1e-10*zero[i]
      plambda[i] <- exp(pmu.count[i])
      pmu.count[i] <-    beta1*x1[i] +beta2*x2[i] + beta3*x3[i]  + beta5*race[i] + beta6*age[i]
      
      ## Zero-Inflation
      pzero[i] ~ dbern(ppi[i])
      ppi[i] <- ilogit(pmu.binary[i])
      pmu.binary[i] <-  alpha1*x1[i] +alpha2*x2[i] + alpha3*x3[i]   + alpha5*race[i] + alpha6*age[i]
    }  

    for (j in 1:J){
     alpha.j[j] ~ dnorm(mu.beta, tau.beta)
    }
  
    for (j in 1:J){
     beta.j[j] ~ dnorm(mu.beta, tau.beta)
    }
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 

    alpha1 ~ dnorm(mu.beta,tau.beta) 
    alpha2 ~ dnorm(mu.beta,tau.beta) 
    alpha3 ~ dnorm(mu.beta,tau.beta) 
    alpha5 ~ dnorm(mu.beta,tau.beta) 
    alpha6 ~ dnorm(mu.beta,tau.beta) 

}"

dat <- data.frame(x=type_of_gun,y=number_of_victims,region = allshootings$region)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                region = dat$region,
                y=dat$y, 
                J = 4,
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6","alpha1","alpha2","alpha3","alpha4","alpha5","alpha6","alpha.j","beta.j")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))
```

When we cluster by region, we maintain the relationship of rifle to number of victims, which is good because it means the observed relationship is not an artifact of the aggregated data.
