---
title: "Bayes Project (Casey's Part)"
author: "Bianca, Heather, Casey"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(rjags)
library(R2jags)
```

## Data

```{r reading in data combining csvs, warning=FALSE, results='hide'}

#setwd("/Users/gcgibson/BayesProject/")

shootingsafter <- read.csv("/Users/gcgibson/BayesProject/ShootingsAfter1991.csv")
moreshootings <- read.csv("/Users/gcgibson/BayesProject/Shootings2016.csv")

write.csv(rbind(shootingsafter,moreshootings), "AllShootings.csv")

allshootings <- read.csv("/Users/gcgibson/BayesProject/AllShootings.csv")

ames.data <- allshootings$Victim.s..Deceased..at.school.

number_of_victims <- c()
for (i in 1:length(ames.data)){
  if (is.na(ames.data[i])){
    #print ("hello")
    number_of_victims <- c(number_of_victims,0)
  } else if (ames.data[i] == "None"){
    number_of_victims <- c(number_of_victims,0)
  } else{
    number_of_victims <- c(number_of_victims,ames.data[i])
  }
}

plot(number_of_victims)


type_of_gun <- matrix(0,nrow=length(number_of_victims),ncol=4)
tmp <- allshootings$Weapon.s..Categories

for (i in 1:length(tmp)){
  if (tmp[i] == "Handgun"){
    type_of_gun[i,1] =  1
  } else if (tmp[i] == "Rifle"){
    type_of_gun[i,2] =  1
  }  else if (tmp[i] == "Shotgun"){
    type_of_gun[i,3] =  1
  }  else {
    type_of_gun[i,4] =  1
  }
}
plot(type_of_gun)
```


We first attempt a simple poisson regression. 

### Poisson


```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
for(i in 1:N){
      y[i] ~ dpois(lambda[i])
      log(lambda[i]) <- mu[i]
      mu[i] <- beta4 + beta1*x1[i] +beta2*x2[i] + beta3*x3[i]    
      }     
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```

We see that neither handgun nor rifle has a significant effect on the number of victims. 

What if we control for race and age?

### Poisson w Covariates

```{r}
race <- allshootings$Shooter.s..or.Attacker.s..Race
race_clean <- c()
for (i in 1:length(race)){
  if (race[i] == "African American"){
    race_clean <- c(race_clean,0)
  } else if (race[i] == "Caucasian"){
    race_clean <- c(race_clean,1)
  }  else if (race[i] == "Hispanic"){
    race_clean <- c(race_clean,2)
  }   else{
    race_clean <- c(race_clean,4)
  }
}
age <- allshootings$Shooter.s..or.Attacker.s..Age
age_clean <- c()
for (i in 1:length(age)){
  if ( 0 < as.numeric(age[i]) & as.numeric(age[i]) <10 ){
    age_clean <- c(age_clean,0)
  } else if (10 < as.numeric(age[i]) & as.numeric(age[i]) < 20){
    age_clean <- c(age_clean,1)
  }   else{
    age_clean <- c(age_clean,2)
  }
}
```




```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
for(i in 1:N){
      y[i] ~ dpois(lambda[i])
      log(lambda[i]) <- mu[i]
      mu[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      }     
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 
}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```


We still don't see any effect of rifle on number of victims, but we see a negative effect of a handgun. That is, we can't say that rifles kill more people than other weapons categories (what I hoped we would find), but we can say that handguns kill fewer people than other weapons. From a policy perspective it makes sense to limit weapons to handguns. 


What happens if we control for the large number of zeros present in the data?

### Zero Inflated Poisson w Covariates

```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
for(i in 1:N){
      y[i] ~ dpois(lambda.hacked[i])
      lambda.hacked[i] <- lambda[i]*(1-zero[i]) + 1e-10*zero[i]
      lambda[i] <- exp(mu.count[i])
      mu.count[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      
      ## Zero-Inflation
      zero[i] ~ dbern(pi[i])
      pi[i] <- ilogit(mu.binary[i])
      mu.binary[i] <- alpha1*x1[i] +alpha2*x2[i] + alpha3*x3[i] + alpha4  + alpha5*race[i] + alpha6*age[i]
    }  
    ## Priors 
  
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 

    alpha1 ~ dnorm(mu.beta,tau.beta) 
    alpha2 ~ dnorm(mu.beta,tau.beta) 
    alpha3 ~ dnorm(mu.beta,tau.beta) 
    alpha4 ~ dnorm(mu.beta,tau.beta) 
    alpha5 ~ dnorm(mu.beta,tau.beta) 
    alpha6 ~ dnorm(mu.beta,tau.beta) 

}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6","alpha1","alpha2","alpha3","alpha4","alpha5","alpha6")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```


Ah-ha! if we use the Zero-inflated model we see that rifles do have a positive association with a higher number of victims. 



Just for fun we throw in an $AR(1)$ error process since our data is time-series data. 

### Poisson + Ar with Covariates

```{r}
library(rjags)
library(R2jags)
model <- "model {
    ## Likelihood
mu.count[1] <-  beta1*x1[1] +beta2*x2[1] + beta3*x3[1] + beta4  + beta5*race[1] + beta6*age[1]
for(i in 2:N){
      y[i] ~ dpois(lambda[i])
      lambda[i] <- exp(mu[i])
      mu[i] <- mu.count[i] + ar1 * ( y[i-1] - mu.count[i-1] )
      mu.count[i] <- beta1*x1[i] +beta2*x2[i] + beta3*x3[i] + beta4  + beta5*race[i] + beta6*age[i]
      
      
    }  
    ## Priors 
    ar1 ~ dunif(-1.1,1.1) 
    beta1 ~ dnorm(mu.beta,tau.beta) 
    beta2 ~ dnorm(mu.beta,tau.beta) 
    beta3 ~ dnorm(mu.beta,tau.beta) 
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    beta6 ~ dnorm(mu.beta,tau.beta) 

 

}"

dat <- data.frame(x=type_of_gun,y=number_of_victims)

forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4=dat$x.4,
                age = age,
                race = race,# predictors
                y=dat$y,  # DV
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=1)  # diffuse priors

parnames <- c( "beta1","beta2","beta3","beta4","beta5","beta6","alpha1","alpha2","alpha3","alpha4","alpha5","alpha6")
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 1000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array
#hist(c(mcmc.array[,,"beta[1]"]), freq = F, main = "", xlab ="Intercept")
#hist(c(mcmc.array[,,"beta[2]"]), freq = F, main = "", xlab ="Slope")
print ("Effect of Handgun")
print (quantile(mcmc.array[,,"beta1"],c(.025,.975)))
print ("Effect of Rifle")
print (quantile(mcmc.array[,,"beta2"],c(.025,.975)))

```



