---
title: "Final Report"
author: "Heather Weaver, Bianca Doone, Casey Gibson"
date: "4/21/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1)
$$Pr(y_i = j ) = \begin{cases}
\pi_i  + (1-\pi_i)g(y_i = 0) & j=0 \\
(1-\pi_i)g(y_i ) & j >0 
\end{cases}$$


$$g \sim Poisson(\lambda_i)$$

where 
$$logit(\pi_{i}) = \alpha_{0j[i]} + \alpha_{1j[i]}x_{1} +\alpha_{2j[i]}x_{2} + \alpha_{3j[i]}x_{3} + \alpha_{4}(race_i-\bar{race}) + \alpha_5(age_i - \bar{age})$$
$$log(\lambda_{i}) = \beta_{0j[i]} + \beta_{1j[i]}x_{1} +\beta_{2j[i]}x_{2} + \beta_{3j[i]}x_{3} + \beta_{4}(race_i-\bar{race}) + \beta_5(age_i - \bar{age})$$



$$\begin{pmatrix}
\beta_{0j[i]} \\
\beta_{1j[i]} \\
\beta_{2j[i]} \\ 
\beta_{4j[i]}
\end{pmatrix} \sim N\begin{bmatrix} \begin{pmatrix}
\mu_{\beta} \\
\mu_{\beta} \\
\mu_{\beta}  \\
\mu_{\beta} 
\end{pmatrix} ,\Sigma
\end{bmatrix}
$$

$$\begin{pmatrix}
\alpha_{0j[i]} \\
\alpha_{1j[i]} \\
\alpha_{2j[i]} \\ 
\alpha_{4j[i]}
\end{pmatrix} \sim N\begin{bmatrix} \begin{pmatrix}
\mu_{\alpha} \\
\mu_{\alpha} \\
\mu_{\alpha}  \\
\mu_{\alpha} 
\end{pmatrix} ,\Gamma
\end{bmatrix}
$$

$$\sigma \sim InvWeishart()$$
$$\Gamma \sim InvWeishart()$$




### 2)

```{r}
library(readr)
library(ggplot2)
library(rjags)
library(R2jags)


#setwd("/Users/gcgibson/BayesProject/")

shootingsafter <- read.csv("/Users/gcgibson/BayesProject/ShootingsAfter1991.csv")
moreshootings <- read.csv("/Users/gcgibson/BayesProject/Shootings2016.csv")


allshootings <- read.csv("/Users/gcgibson/BayesProject/AllShootings_clean_final.csv")

#gunRegulations <- read.csv("/Users/gcgibson/BayesProject/Gu")

ames.data <- allshootings$Victim.s..Injured..at.school.

number_of_victims <-c()
s <- c()
for (i in 1:length(ames.data)){
  if (is.na(ames.data[i])){
    #print ("hello")
    number_of_victims <- c(number_of_victims,0)
  } else if (ames.data[i] == "None"){
    number_of_victims <- c(number_of_victims,0)
  } else{
    number_of_victims <- c(number_of_victims,ames.data[i])
  }
}

allshootings$number_of_victims <- number_of_victims

#plot(1:length(number_of_victims),number_of_victims,ylab="Number #of Victims",xlab="Shooting ID ordered by time")
p <- ggplot(allshootings, aes(x=seq(1,1113),y = allshootings$number_of_victims)) +  geom_point() + labs(title = "Number of Victims in each shooting incident from 1990-2016", y="Number of Victims", x="1990-2016")   + scale_x_discrete(breaks = c(1,500,1100), labels=c("1990","2002","2016")) +theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(p)

type_of_gun <- matrix(0,nrow=length(number_of_victims),ncol=4)
tmp <- allshootings$Weapon.s..Categories

for (i in 1:length(tmp)){
  if (tmp[i] == "Handgun"){
    type_of_gun[i,1] =  1
  } else if (tmp[i] == "Rifle"){
    type_of_gun[i,2] =  1
  }  else if (tmp[i] == "Shotgun"){
    type_of_gun[i,3] =  1
  }  else {
    type_of_gun[i,4] =  1
  }
}

race <- allshootings$Shooter.s..or.Attacker.s..Race
race_clean <- c()
for (i in 1:length(race)){
  if (race[i] == "African American"){
    race_clean <- c(race_clean,0)
  } else if (race[i] == "Caucasian"){
    race_clean <- c(race_clean,1)
  }  else if (race[i] == "Hispanic"){
    race_clean <- c(race_clean,2)
  }   else{
    race_clean <- c(race_clean,4)
  }
}
age <- allshootings$Shooter.s..or.Attacker.s..Age
age_clean <- c()
for (i in 1:length(age)){
  if ( 0 < as.numeric(age[i]) & as.numeric(age[i]) <10 ){
    age_clean <- c(age_clean,0)
  } else if (10 < as.numeric(age[i]) & as.numeric(age[i]) < 20){
    age_clean <- c(age_clean,1)
  }   else{
    age_clean <- c(age_clean,2)
  }
}

dat <- data.frame(x=type_of_gun,y=number_of_victims,race=race_clean -mean(race_clean),age=age_clean-mean(age_clean))

ones_ <- c()
for (y in dat$y){
  if (y == 0){
    ones_ <- c(ones_,0)
  }
  else{
    ones_<-c(ones_,1)
  }
}

y_fact <- c()
for (y in dat$y){
  
    y_fact <- c(y_fact,factorial(y))
 
}



allshootings$region[  allshootings$State == "ME"| 
                allshootings$State == "MA"| 
                allshootings$State == "NH"| 
                allshootings$State == "RI"| 
                allshootings$State == "VT" ] <- 5

allshootings$region[allshootings$State == "CT" | 
                allshootings$State == "NJ"| 
                allshootings$State == "NY"| 
                allshootings$State == "PA"|
                  allshootings$State == "WV"] <- 1

allshootings$region[allshootings$State == "IL" | 
                allshootings$State == "IN"| 
                allshootings$State == "MI"| 
                allshootings$State == "OH"| 
                allshootings$State == "WI"] <-6 
                  
allshootings$region[allshootings$State == "MN"| 
                allshootings$State == "MO"| 
                allshootings$State == "NE"| 
                allshootings$State == "ND"| 
                allshootings$State == "SD"|
                 allshootings$State == "KS"|
                  allshootings$State == "IA"] <- 2

allshootings$region[allshootings$State == "DE" | 
                allshootings$State == "GA"| 
                allshootings$State == "MD"| 
                allshootings$State == "NC"| 
                allshootings$State == "SC"| 
                allshootings$State == "VA"| 
                allshootings$State == "FL"| 
                allshootings$State == "Wash D.C" ] <- 7

allshootings$region[allshootings$State == "AL"| 
                allshootings$State == "KY"| 
                allshootings$State == "MS"| 
                allshootings$State == "TN"| 
                allshootings$State == "AR"| 
                allshootings$State == "LA"| 
                allshootings$State == "OK"| 
                allshootings$State == "TX"] <- 3

allshootings$region[allshootings$State == "AZ" | 
                allshootings$State == "CO"| 
                allshootings$State == "ID"| 
                allshootings$State == "MT"| 
                allshootings$State == "NV"| 
                allshootings$State == "NM"| 
                allshootings$State == "UT" |
                  allshootings$State == "WY"] <- 8 


allshootings$region[allshootings$State == "AK"| 
                allshootings$State == "CA"| 
                allshootings$State == "HI"| 
                allshootings$State == "OR"| 
                allshootings$State == "WA"] <- 4


gun_reg <- read.csv("/Users/gcgibson/BayesProject/Gun_Reg.csv")

region_1_gun_reg <- sum(gun_reg[gun_reg$State == "Connecticut",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "New Jersey",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "New York",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "Pennsylvania",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "West Virginia",][1,3:ncol(gun_reg)])




region_2_gun_reg <- sum(gun_reg[gun_reg$State == "Minnesota",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Nebraska",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Missouri",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "North Dakota",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "South Dakota",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Kansas",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Iowa",][1,3:ncol(gun_reg)])




region_3_gun_reg <- sum(gun_reg[gun_reg$State == "Alabama",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Kentuky",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Mississippi",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Tennessee",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "Arkansas",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "Louisiana",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "Oklahoma",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "Texas",][1,3:ncol(gun_reg)])






 region_4_gun_reg <- sum(gun_reg[gun_reg$State == "Alaska",][1,3:ncol(gun_reg)]) 
 + sum(gun_reg[gun_reg$State == "California",][1,3:ncol(gun_reg)])
 + sum(gun_reg[gun_reg$State == "Hawaii",][1,3:ncol(gun_reg)])
  + sum(gun_reg[gun_reg$State == "Oregon",][1,3:ncol(gun_reg)])
  + sum(gun_reg[gun_reg$State == "Washington",][1,3:ncol(gun_reg)])




region_5_gun_reg <- sum(gun_reg[gun_reg$State == "Maine",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "Massachusetts",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "New Hampshire",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "Rhode Island",][1,3:ncol(gun_reg)])
+ sum(gun_reg[gun_reg$State == "Vermont",][1,3:ncol(gun_reg)])



region_6_gun_reg <- sum(gun_reg[gun_reg$State == "Illonois",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "Michigan",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "Ohio",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "Wisconsin",][1,3:ncol(gun_reg)]) 
+ sum(gun_reg[gun_reg$State == "Indiana",][1,3:ncol(gun_reg)]) 




region_7_gun_reg <-  sum(gun_reg[gun_reg$State == "Delaware",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Florida",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Georgia",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Maryland",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "North Carolina",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "South Carolina",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Virginia",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Wash D.C",][1,3:ncol(gun_reg)]) 



region_8_gun_reg <-  sum(gun_reg[gun_reg$State == "Arizona",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Colorado",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Idaho",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Montana",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "New Mexico",][1,3:ncol(gun_reg)])
+sum(gun_reg[gun_reg$State == "Utah",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Nevada",][1,3:ncol(gun_reg)]) 
+sum(gun_reg[gun_reg$State == "Wyoming",][1,3:ncol(gun_reg)]) 


gun_reg_vec <- c(region_1_gun_reg,region_2_gun_reg,region_3_gun_reg,region_4_gun_reg,region_5_gun_reg,region_5_gun_reg,region_7_gun_reg,region_8_gun_reg,0)


model <- "model{

    # reference for C and zero trick http://www.medicine.mcgill.ca/epidemiology/Joseph/courses/common/Tricks.html
    C <- 10000 

    for(j in 1:ntrain){
      LikCountModel[j] <- pow(mu[j],y[getitrain.j[j]])/y_fact[getitrain.j[j]]*exp(-mu[j])
      
    # beta part
      eta[j] <- beta.j[1,region[j]] + beta.j[2,region[j]]*x2[getitrain.j[j]] + beta.j[3,region[j]]*x3[getitrain.j[j]]   +   beta.j[4,region[j]]*x4[getitrain.j[j]] + beta4*race[getitrain.j[j]] + beta5*age[getitrain.j[j]]
      
      mu[j] <- exp(eta[j])
      #alpha part
     zeta[j] <- alpha.j[1,region[j]] + alpha.j[2,region[j]]*x2[getitrain.j[j]] + alpha.j[3,region[j]]*x3[getitrain.j[j]]   +   alpha.j[4,region[j]]*x4[getitrain.j[j]] + alpha4*race[getitrain.j[j]] + alpha5*age[getitrain.j[j]]
      w[j] <- exp(zeta[j])/(1+exp(zeta[j]))
      
      #1/0 Tricks: ones is a column containing only ones, with the same size of y
      p[j] <- L[j] / C
      ones[j] ~ dbern(p[j])
      
      #Full likelihood expression
      L[j] <- LikCountModel[j] * (1-w[j]) + equals(y[j],0)*w[j]
      
    } 
  
    for (j in 1:4){
      mu_beta_vec[j] <- mu_beta 
    }

    for (j in 1:4){
      mu_alpha_vec[j] <- mu_alpha 
    }
    
    for (j in 1:J){
       beta.j[1:4,j] ~ dmnorm( mu_beta_vec, V)
     
    }
  
     for (j in 1:J){
       alpha.j[1:4,j] ~ dmnorm( mu_alpha_vec, W)
     }
  
    V ~ dwish(R, 4)
    W ~ dwish(R, 4)
    sigma ~ dunif(0,20)
    alpha4 ~ dnorm(mu.alpha,tau.alpha)
    alpha5 ~ dnorm(mu.alpha,tau.alpha)
    beta4 ~ dnorm(mu.beta,tau.beta) 
    beta5 ~ dnorm(mu.beta,tau.beta) 
    mu_beta ~ dnorm(0,.01)
    mu_alpha ~ dnorm(0,.01)
    
    r ~ dunif(0,50)
}"



dat$region <- allshootings$region

dat$region[is.na(dat$region)] <- 9
forJags <- list(x1=dat$x.1,
                x2=dat$x.2,
                x3=dat$x.3,
                x4 = dat$x.4,
                race = dat$race ,
                age = dat$age,
                y=dat$y,
                ones = ones_,# DV
                y_fact = y_fact,
                R = diag(c(1,1,1,1)),
                gun_reg_vec = gun_reg_vec,
                prec_vec = solve(diag(4)),
                region = dat$region,
                J = 9,
                N=1113,  # sample size
                mu.beta=0,  # priors centered on 0
                tau.beta=.01,
                mu.alpha=0,  # priors centered on 0
                tau.alpha=.01,
                mu.gamma = 0,
                tau.gamma=.01)  # diffuse priors


parnames <- c( "beta.j","mu","beta4","beta5")

train_index <- sample(1:1113,890)
test_index <- setdiff(1:1113,train_index)


forJags$getitrain.j <- train_index
forJags$getitest.j <- test_index
forJags$ntrain <- length(train_index)



mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 5000, n.thin = 10, model.file = textConnection(model))

mcmc.array <- mod$BUGSoutput$sims.array

library(shinystan)
#my_sso <- launch_shinystan(as.shinystan(mod$BUGSoutput$sims.array))



print ("Regional effect of rifle ")
for (j in 1:8){
  print (quantile(mod$BUGSoutput$sims.list$beta.j[,2,j],c(.025,.975)))

}
print (quantile(mod$BUGSoutput$sims.array[,,"gamma0"]))

```



Since we can't explicity encode yrep in the model, we have to use the samples of the parameters of the training set. For simplicity, we focus only on the first region. Technically, we should also include the alphas, but to keep things simple we will consider posterior-predictive only based on the betas. 



```{r}
beta_star_j <- mod$BUGSoutput$sims.list$beta.j[,,1]

beta_star_0 <- beta_star_j[,1]
beta_star_1 <- beta_star_j[,2]
beta_star_2 <- beta_star_j[,3]
beta_star_3 <- beta_star_j[,4]

beta_star_4 <- mod$BUGSoutput$sims.array[,,"beta4"]
beta_star_5 <- mod$BUGSoutput$sims.array[,,"beta5"]


yrep_zinb <- matrix(NA,nrow = length(test_index),ncol=100)
count <- 1
for (i in test_index){
  tmp <- c()
  for (mc_sample in 1:100){
    tmp <- c(tmp,rpois(1,exp(beta_star_0[mc_sample] + beta_star_1[mc_sample]*dat$x.2[i] + beta_star_2[mc_sample]*dat$x.3[i] +beta_star_3[mc_sample]*dat$x.4[i] + beta_star_4[mc_sample]*dat$race[i] + beta_star_5[mc_sample]*dat$age[i] )))
  }
  yrep_zinb[count,] <- tmp
  count <- count +1
}

plot(test_index,rowMeans(yrep_zinb) - dat$y[test_index],ylab="y_hat_i - y_i",xlab="Test Index",main="Out of sample predictions")


```


We can now examine the coverage probability. We do this by asking if the true observation falls within the 95% credible interval given by the monte carlo samples.

```{r}
correct <- 0
for (i in 1:nrow(yrep_zinb)){
  true_y <- dat$y[train_index[i]]
  if (true_y >= quantile(yrep_zinb[i,],.025) && true_y <= quantile(yrep_zinb[i,],.975)){
    correct <- correct + 1
  }
  
}

print ("Empirical coverage probability")
print (correct/nrow(yrep_zinb))
```

Pretty low coverage probability but I'm not sure how to handle the fact that this is a discrete distribution. 


### 3

Lets examine what happens when we instead leave out observations not at random, but according to time. Suppose we leave out the first 20% of observations


```{r}
train_index <- 223:1113
test_index <- 1:223


forJags$getitrain.j <- train_index
forJags$getitest.j <- test_index
forJags$ntrain <- length(train_index)
mod <- jags(data = forJags, 
                    parameters.to.save=parnames, 
                    n.chains = 3, n.burnin = 1500, n.iter =1500 + 5000, n.thin = 10, model.file = textConnection(model))
```


```{r}
beta_star_j <- mod$BUGSoutput$sims.list$beta.j[,,1]

beta_star_0 <- beta_star_j[,1]
beta_star_1 <- beta_star_j[,2]
beta_star_2 <- beta_star_j[,3]
beta_star_3 <- beta_star_j[,4]

beta_star_4 <- mod$BUGSoutput$sims.array[,,"beta4"]
beta_star_5 <- mod$BUGSoutput$sims.array[,,"beta5"]


yrep_zinb <- matrix(NA,nrow = length(test_index),ncol=100)
count <- 1
for (i in test_index){
  tmp <- c()
  for (mc_sample in 1:100){
    tmp <- c(tmp,rpois(1,exp(beta_star_0[mc_sample] + beta_star_1[mc_sample]*dat$x.2[i] + beta_star_2[mc_sample]*dat$x.3[i] +beta_star_3[mc_sample]*dat$x.4[i] + beta_star_4[mc_sample]*dat$race[i] + beta_star_5[mc_sample]*dat$age[i] )))
  }
  yrep_zinb[count,] <- tmp
  count <- count +1
}

plot(test_index,rowMeans(yrep_zinb) - dat$y[test_index],ylab="y_hat_i - y_i",xlab="Test Index",main="Out of sample predictions")


correct <- 0
for (i in 1:nrow(yrep_zinb)){
  true_y <- dat$y[train_index[i]]
  if (true_y >= quantile(yrep_zinb[i,],.025) && true_y <= quantile(yrep_zinb[i,],.975)){
    correct <- correct + 1
  }
  
}

print ("Empirical coverage probability")
print (correct/nrow(yrep_zinb))
```



```{r}
loglike1.model2<- mod2$BUGSoutput$sims.list$L
loo2 <- loo(loglike1.model2)
print (loo2$elpd_loo)



psis2 <- psislw(-loglike1.model2)
lw2 <- psis2$lw_smooth
ppc_loo_pit_overlay(
  y = dat$y,
  yrep = t(yrep.i_zinb),
  lw = lw2
)
```

