---
title: 'Heather''s Part '
author: "Heather"
date: "3/5/2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3.5, fig.width = 5, fig.align = "center")
library(readr)
library(ggplot2)
library(googledrive)
library(rjags)
library(R2jags)
library(dplyr)

PlotTrace <- function(parname, mcmc.array,n.chains= NULL, n.sim= NULL, main = NULL){
  if (is.null(main)) main <- parname
  if (is.null(n.sim)) n.sim <- dim(mcmc.array)[1]
  if (is.null(n.chains)) n.chains <- dim(mcmc.array)[2]
  plot(c(mcmc.array[,1,parname]), type = "l", ylab = main,  main = main,
       ylim = c(min(mcmc.array[,,parname]),max(mcmc.array[,,parname])))
  for (chain in 1:n.chains){
    lines(c(mcmc.array[,chain,parname]), type = "l", col = chain)
  }
  for (chain in 1:n.chains){
    curve(predict(loess(c(mcmc.array[,chain,parname])~seq(1,n.sim)),x), lty = 2, lwd = 3, add = TRUE, type = "l", col = chain)
  }
}

allshootings <- read.csv("~/AllShootings_clean.csv")
gun_regulations <- read.csv("~/gun_regulation.csv")
```

Soem exploratory/summary graphs:  

```{r plots, echo=FALSE, warning = "hide"}
dates <- allshootings$Date
dates <- dates[2:length(dates)]

allshootings$Date <- as.Date(allshootings$Date,format="%m/%d/%Y")
allshootings$year <- format(allshootings$Date,"%Y")

per_state <- matrix(nrow=53, ncol=2)
state.i <- as.vector(allshootings$State)
state.k <- unique(state.i)
K <- length(state.k)
number_per_state <- rep(NA, K)
for (i in 1:length(summary(allshootings$State))){
  per_state[i,2] <- summary(allshootings$State)[i]
  per_state[i,1] <- sort(state.k)[i]
}
state <- as.data.frame(per_state)

number <- rep(NA, length(allshootings$Victim.s..Deceased..at.school.))
for (i in 1:length(allshootings$Victim.s..Deceased..at.school.)){
  if (is.na(allshootings$Victim.s..Deceased..at.school.)[i]==TRUE){
    number[i] <- 0
  }
    else{
      number[i]<- allshootings$Victim.s..Deceased..at.school.[i]
  }
}

allshootings$Victim.s..Deceased..at.school. <- number

number2 <- rep(NA, length(allshootings$Victim.s..Deceased..at.school.))

for (i in 1:length(allshootings$Victim.s..Deceased..off.campus.)){
  if (is.na(allshootings$Victim.s..Deceased..off.campus.)[i]==TRUE){
    number2[i] <- 0
  }
    else{
      number2[i]<- allshootings$Victim.s..Deceased..off.campus.[i]
  }
}

allshootings$Victim.s..Deceased..off.campus. <- number2

allshootings$total_victims <- allshootings$Victim.s..Deceased..at.school. + allshootings$Victim.s..Deceased..off.campus.


ggplot(gun_regulations, aes(x=gun_regulations$state,y = gun_regulations$lawtotal)) + geom_point()   + labs(title = "Total Number of Gun Regulations by State")

ggplot(allshootings,aes(x=allshootings$Weapon.s..Categories,y = allshootings$total_victims)) + geom_point()   + labs(title = "Number of Victims by Gun Type", y="Number of Victims", x="Type of Gun")

#Want to have differnt colors on the plot below for states with high gun regulations(>50) vs. states with low regulations. 
ggplot(state,aes(x=state$V1,y = state$V2)) + geom_point()  + labs(title = "Number of Incidents by State", y="Number of Incidents", x="State")
```
Goal: Model gun type with gun regualtions to see if increased gun regulations play a role in type of gun used in school shootings. 

Still to do:   
 - Get code below to run for simple model and model with gun regulations.  
 - Check that priors for all models are vague. (histograms of priors and posteriors)  
 - Check for converenge issues (traceplots, Rhat and neff for all parameters)



Simple model for gun type used: 

$$ \text{ gun type } = G_{j} \sim Multinomial(N, p_1, ..., p_J) $$
$$log(\frac {P(G_j = j|g.reg_i)}{P(G_j = J|g.reg_i)}) = \alpha_j   $$
$$ j = 1, ..., J-1 $$
$$ J \text{ is the reference gun type} $$

```{r, eval=FALSE}
group.i <- as.vector(allshootings$Weapon.s..Categories)
n <- length(allshootings$Weapon.s..Categories)
group.j <- unique(group.i)
J <- length(group.j)
weapongetj.i <- rep(NA, n)
for (j in 1:J){
  weapongetj.i[group.i==group.j[j]] <- j
}

model <- "
model {
weapon_type[1:J] ~ dmultinom(p[1:J],N)

for (j in 1:J) {
              p[j] <- delta[j] / sum(delta[j])
              delta[j] ~ dgamma(alpha[j], 0.0001)
}

for (j in 1 : J) { 
  alpha[j] ~ dunif(0.001, .8)
  
}

} # end model
"
jags.data <- list(weapon_type = allshootings$Weapon.s..Categories, J=J,
                  N=length(allshootings$Weapon.s..Categories))

parnames <- c( "alpha")

mod <- jags(data = jags.data, parameters.to.save=parnames, n.chains = 3, n.burnin = 1000, n.iter = 2500+1000, model.file = textConnection(model))
```
Using gun regulations to model gun type used: 

$$ \text{ gun type } = G_{j} \sim Multinomial(N, p_1, ..., p_J) $$
$$log(\frac {P(G_j = j|g.reg_i)}{P(G_j = J|g.reg_i)}) = \alpha_{j} + \beta_{j}*g.reg_{s,t[i]}  $$
$$ where gun regulations is a binary variable, either high or low in state, s, at time, t.
```{r, eval=FALSE}
model <- "
model {
weapon_type[1:J] ~ dmultinom(p[1:J], N)

for (j in 1:J) {
              p[j] <- delta[j] / sum(delta[j])
              delta[j] ~ dgamma(phi[j], 0.0001)
}
for (j in 1 : J) { 
  phi[j] <- alpha[j] 	+ beta*gun_reg
  
}

for (j in 1 : J) { 
  alpha[j] ~ dunif(0.001, .8)
  
}

  beta ~ rnorm(0, 0.0001)


} # end model
"
jags.data <- list(weapon_type = allshootings$Weapon.s..Categories, J=J, 
                  gun_reg = gun_regulations$lawtotal,
                  n=length(allshootings$Weapon.s..Categories))

parnames <- c( "alpha", "beta")

mod <- jags(data = jags.data, parameters.to.save=parnames, 
            n.chains = 3, n.burnin = 1000, n.iter = 2500, 
            model.file = textConnection(model))
```



